---
title: "Regression models"
subtitle: "Notes"
author: "Pablo Benavides-Herrera"
date: 2020-07-06
output: 
  html_notebook:
    toc: TRUE
    toc_float: TRUE
    highlight: tango
    theme: readable
    number_sections: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
source("regression_code.R", local = knitr::knit_global())
```

```{r pkgs}
library(DT)
```

# US % change data 

We are using the `us_change` dataset, available from the `fpp3` package.

```{r}
datatable(us_change)
```

## Exploratory data analysis

### Time plot

We begin by plotting each variable in a time plot.

```{r}
time_plots
```

### Pairs plot

```{r}
pairs
```

Withe the pairs plot, we can see:

- the scatterplots for every pair of variables, 
- each variable's distribution 
- and their correlation. 


## Multiple linear regression {.smaller}

```{r}
report(cons_fit)
```

<div class="notes">
The multiple linear regression includes all available variables, with an $R^2 = 0.7683$.
</div>

## Residuals diagnostics

### Data vs. fitted values

```{r}
cons_data_fit
```

<div class="notes">
The fit is looking good.
</div>

```{r}
cons_data_fit2
```

<div class="notes">
We see that the residuals are looking better now.
</div>

### Residuals vs. predictors

```{r}
cons_resid_predictors
```

<div class="notes">
The residuals plotted against each predictor show a random pattern, which is a sign of a good fit.
</div>

### Residuals vs. fitted values

```{r}
cons_resid_fitted
```

<div class="notes">
The residuals vs. the fitted values show a random pattern too. It seems that our model is a good one.
</div>


```{r}
cons_fit %>% 
  gg_tsresiduals()
```

