---
title: "Regression models"
author: "Pablo Benavides-Herrera"
date: 2020-07-06
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
source("regression_code.R", local = knitr::knit_global())
library(DT)
```

## US % change data {.smaller}

```{r}
datatable(us_change)
```

## Time plots

```{r}
time_plots
```


## Pairs plot

```{r}
pairs
```

<div class="notes">
We can see 

- the scatterplots for every pair of variables, 
- each variable's distribution 
- and their correlation. 
</div>

## Multiple linear regression {.smaller}

```{r}
report(cons_fit)
```

<div class="notes">
The multiple linear regression includes all available variables, with an $R^2 = 0.7683$.
</div>

# Residuals diagnostics

## Data vs. fitted values

```{r}
cons_data_fit
```

<div class="notes">
The fit is looking good.
</div>

## 

```{r}
cons_data_fit2
```

<div class="notes">
We see that the residuals are looking better now.
</div>

## Residuals vs. predictors

```{r}
cons_resid_predictors
```

<div class="notes">
The residuals plotted against each predictor show a random pattern, which is a sign of a good fit.
</div>

## Residuals vs. fitted values

```{r}
cons_resid_fitted
```

<div class="notes">
The residuals vs. the fitted values show a random pattern too. It seems that our model is a good one.
</div>

## 

```{r}
cons_fit %>% 
  gg_tsresiduals()
```

