---
title: "`R` Tutorial"
author: "Pablo Benavides-Herrera"
date: 2020-08-20
tutorial:
  id: "r_101"
  version: 0.1
output: 
  learnr::tutorial:
    progressive: TRUE
    allow_skip: TRUE
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
knitr::opts_chunk$set(echo = FALSE)
```


## Calentamiento

```{r quiz-warmup-q}
quiz(caption = "Veamos si quedó claro lo visto en clase. Pueden utilizar el espacio de abajo para probar con código.",
     question("¿Cuál es el símbolo para agregar comentarios en un R Script?", 
              answer("#", 
                     message = "Exacto, se requiere escribir un `#` al inicio del renglón para que `R` sepa que lo que sigue no debe ejecutarlo.", 
                     correct = TRUE),
              answer("<", 
                     message = "Esto marcaría un error."),
              answer(" %",
                     message = "Así se agregaría un comentario en `MATLAB`, no en `R`."),
              answer("$",
                     message = "Esto marcaría un error."),
              random_answer_order = TRUE,
              allow_retry = TRUE
              ),
     question("¿Para qué sirve agregar secciones en un R script? (el atajo del teclado es CTRL *(o CMD en Mac)* + SHIFT + R)",
              answer("Para mostrar/ocultar partes del código.",
                     correct = TRUE),
              answer("Para poder desplazarse rápidamente entre las partes del código.",
                     correct = TRUE),
              answer("Para poder crear sub archivos de R."),
              answer("Para poder observar una guía con el nombre de cada sección y navegar como si fuera un índice.",
                     correct = TRUE),
              random_answer_order = TRUE,
              allow_retry = TRUE
              ),
     question("¿De qué maneras se pueden correr una o más filas de código de un R script?",
              answer("Colocando el cursor en la fila deseada y oprimir el botón de **'RUN'**.",
                     correct = TRUE),
              answer("Colocando el cursor en la fila deseada y presionar **CTRL + Enter**",
                     correct = TRUE),
              answer("Colocando el cursor en la consola y oprimir CTRL + ALT + Enter"),
              answer("Seleccionando todas las filas que se deseen correr y oprimir 'RUN' o CTRL + Enter.",
                     correct = TRUE)
              )
     )
```

```{r quiz-warmup-ex, exercise=TRUE, exercise.lines=10}

```


## Ejercicios

*Complete con el código necesario para responder a cada pregunta*

### Instalación y carga de paqueterías

```{r install-pkgs}
question("¿Cuál es la función que nos permite instalar paqueterías que se encuentren disponibles en el CRAN?",
         answer("`install.packages()`",
                correct = TRUE),
         answer("`install.libraries()`"),
         answer("`install_packages()`"),
         answer("`install_libraries()`"),
         answer("`install.package()`"),
         answer("`install_package()`"),
         incorrect = "Esa función no existe.",
         random_answer_order = TRUE,
         allow_retry = TRUE
         )
```

Carga la paquetería `tidyverse` en el entorno.

```{r lib-tidyverse, exercise=TRUE}

```

```{r lib-tidyverse-solution}
library(tidyverse)
```

### Exercise with Code

*Here's an exercise with some prepopulated code as well as `exercise.lines = 5` to provide a bit more initial room to work.*

Now write a function that adds any two numbers and then call it:

```{r add-function, exercise=TRUE, exercise.lines = 5}
add <- function() {
  
}
```

## Topic 2

### Exercise with Hint

*Here's an exercise where the chunk is pre-evaulated via the `exercise.eval` option (so the user can see the default output we'd like them to customize). We also add a "hint" to the correct solution via the chunk immediate below labeled `print-limit-hint`.*

Modify the following code to limit the number of rows printed to 5:

```{r print-limit, exercise=TRUE, exercise.eval=TRUE}
mtcars
```

```{r print-limit-hint}
head(mtcars)
```

### Quiz

*You can include any number of single or multiple choice questions as a quiz. Use the `question` function to define a question and the `quiz` function for grouping multiple questions together.*

Some questions to verify that you understand the purposes of various base and recommended R packages:

```{r quiz}
quiz(
  question("Which package contains functions for installing other R packages?",
    answer("base"),
    answer("tools"),
    answer("utils", correct = TRUE),
    answer("codetools")
  ),
  question("Which of the R packages listed below are used to create plots?",
    answer("lattice", correct = TRUE),
    answer("tools"),
    answer("stats"),
    answer("grid", correct = TRUE)
  )
)
```

