---
title: "Time series analysis using `fable` & `modeltime`"
subtitle: "Case study: Mobility trends in Mexico in the COVID-19 pandemic"
author: "Pablo Benavides-Herrera"
date: 2020-09-18
output: 
  slidy_presentation:
    incremental: FALSE
    highlight: tango
    fig_width: 7
    fig_height: 6
    df_print: paged
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
```

```{r pkgs}
# tidyverse
library(tidyverse)
library(lubridate)

# plot helpers
library(patchwork)
library(plotly)

# shiny
library(shiny)
library(shinyWidgets)

# tidyverts
library(tsibble)
library(feasts)
library(fable)
library(fable.prophet)

# modeltime
library(tidymodels)
library(timetk)
library(modeltime)
```


## The data

* The dataset is provided by [Google](https://www.google.com/covid19/mobility/) and consists of daily data regarding mobility in Mexico and each of its 32 states.

* It provides time series showing the percentage change in visits to six different type of places, including:

  * Grocery & pharmacy
  * Parks
  * Transit stations
  * Retail & recreation
  * Residential
  * Workplaces
  
* Changes for each day are compared to a baseline value, which is the median for the corresponding day of the week during the period of Jan 3- Feb 6.

## The data (2)

```{r}
df <- read_csv("google_mobility.csv") %>% 
  select(sub_region_1, date:residential_percent_change_from_baseline) %>% 
  pivot_longer(cols = retail_and_recreation_percent_change_from_baseline:residential_percent_change_from_baseline) %>% 
  rename(State = sub_region_1) %>% 
  mutate(name = str_replace(name, "_percent_change_from_baseline",""),
         State = if_else(is.na(State),"Mexico",State),
         State = as_factor(State)) %>% 
  pivot_wider(names_from = name, values_from = value)

renderDataTable(df)

df <- df %>% 
  pivot_longer(cols = -c(State, date),
               names_to = "Location",
               values_to = "Percent_change") %>% 
  mutate(Location = as_factor(Location),
         date = dmy(date),
         Percent_change = Percent_change/100,
         weekday = wday(date, label = TRUE),
         covid = date >= ymd("2020-03-18")) 

df_ts <- df %>% as_tsibble(index = date, key = c(State, Location))
```



## Mobility trends

```{r}
checkboxGroupButtons(inputId = "loc",
                     label = "Type of locations:",
                     choices = levels(df$Location), 
                     selected = "workplaces",
                     justified = TRUE,
                     checkIcon = list(yes = icon("ok", lib = "glyphicon"), no = icon("remove", lib = "glyphicon")))

pickerInput(inputId = "state",
            label = "choose the state(s) to plot",
            choices = levels(df$State),
            selected = "Mexico", 
            multiple = TRUE,
            options = list(`liveSearch` = TRUE,
                           `actions-box` = TRUE)
            )


renderPlotly({
  df %>% 
    filter(State %in% input$state,
           Location %in% input$loc) %>% 
    ggplot(aes(x = date, y = Percent_change, color = State)) +
    geom_line(size = 1) + 
    geom_hline(yintercept = 0, 
               color = "firebrick", 
               linetype = "dashed") + 
    annotate("text", 
             x = ymd("2020-03-28"), 
             y = 0.05,
             label = "COVID-19",
             color = "limegreen") +
    annotate("text", 
             x = last(df$date), 
             y = 0.02,
             label = "Baseline",
             color =  "firebrick") +
    geom_vline(xintercept = as.numeric(ymd("2020-03-18")), 
               linetype = "dotted", 
               size = 0.5, 
               color = "limegreen") +
    facet_wrap(~ Location) + 
    xlab("") + ylab("% change from baseline") +
    scale_y_continuous(labels = scales::percent) +
    theme(legend.position = "none")
})
```


## Exploratory Data Analysis (EDA)

The data exhibit:

* a strong (*weekly*) seasonal pattern, 

* a structural change around March 18th due to the COVID-19 pandemic.

* an upward trend towards the baseline for all the series, except the residential, which has the inverse behavior.

```{r}
dropdownButton(
  tags$h3("Customization"),
  selectInput(inputId = "state2",
            label = "choose the state(s) to plot",
            choices = levels(df$State),
            selected = "Mexico", 
            multiple = TRUE
            ),
  selectInput(inputId = "loc2",
            label = "Type of locations:",
            choices = levels(df$Location), 
            selected = "workplaces", 
            multiple = TRUE
            ),
  sliderTextInput(
  inputId = "date_covid",
  label = "Make calculations based on dates:",
  choices = c("Everything", "Before COVID-19", "During COVID-19"),
  grid = TRUE,
  force_edges = TRUE,
  animate = TRUE
  )
  
  
)


df_eda <- reactive({
  if (input$date_covid == "Everything") {
    df_e <- df_ts
  } else if (input$date_covid == "Before COVID-19"){
    df_e <-  df_ts %>% 
      filter(covid == FALSE)
  } else {
    df_e <- df_ts %>% 
      filter(covid == TRUE) 
  }
  
  df_e %>% 
    filter(State %in% input$state2,
           Location %in% input$loc2)
})

renderDataTable({
  df_eda() %>% 
    features(Percent_change, quantile, prob = seq(0,1, by = 0.25))
})

renderPlotly({
  df_eda() %>%
    ggplot(aes(x = weekday,
               y = Percent_change,
               fill = covid)) +
    geom_boxplot() +
    facet_wrap(~ State) +
    scale_y_continuous(labels = scales::percent)
})

renderPlot({
  df_eda() %>% 
    gg_season(period = "1 week") +
    scale_y_continuous(labels = scales::percent) +
    xlab("") + ylab("% change from baseline") +
    scale_color_viridis_c()
})
```


## The models

We will model this time series using both classical and more advanced models, such as ETS, ARIMA, Seasonal decomposition, linear regressions, facebook's prophet.

```{r}
fable_train <- df_ts %>% 
  filter(State == "Jalisco") %>% 
  slice_head(n = nrow(df_ts) - 15)

fable_test <- df_ts %>% 
  filter(State == "Jalisco") %>% 
  slice_tail(n = 15)

fable_models <- fable_train %>% 
  model(
    ETS = ETS(Percent_change),
    ARIMA = ARIMA(Percent_change),
    `STL + ARIMA` = decomposition_model(
      STL(Percent_change), ARIMA(season_adjust ~ PDQ(0,0,0))),
    LM = TSLM(Percent_change ~ trend() + season()),
    `Piecewise Reg.` = TSLM(Percent_change ~ trend(knots = c(ymd("2020-03-18"),
                                                             ymd("2020-05-03")))),
    Prophet = prophet(Percent_change)
  )
```

The models' accuracy:

```{r}
renderDataTable({
  fable_models %>% 
  fabletools::accuracy()
}) 
```

```{r}
tags$h3("Residual diagnostics")

selectInput(inputId = "fable_mod",
            label = "Choose the model", 
            choices = names(fable_models)[-c(1:2)])
radioButtons(inputId = "fable_series",
             label = "Choose the TS",
             choices = levels(df$Location))

renderPlot({
  fable_models %>% 
  filter(Location == input$fable_series) %>% 
    select(State, Location, input$fable_mod) %>% 
    gg_tsresiduals()
})
```

## Forecasting

```{r}
fable_fcst <- fable_models %>% 
  forecast(h = 15)
```

```{r, fig.height= 12}
renderPlot({
  fable_fcst %>% 
    filter(.model == "ARIMA") %>% 
  autoplot(df_ts)
}) 
```


## modeltime

Using the package `modeltime` we fit an ARIMA, a Boosted ARIMA, ETS, Prophet and linear regression models.

```{r}
jalisco_work <- df %>% 
  filter(State == "Jalisco",
         Location == "workplaces")

splits <- initial_time_split(jalisco_work, prop = 0.9)
```


```{r}
model_fit_arima_no_boost <- arima_reg() %>%
    set_engine(engine = "auto_arima") %>%
    fit(Percent_change ~ date, data = training(splits))

model_fit_arima_boosted <- arima_boost(
    min_n = 2,
    learn_rate = 0.015
) %>%
    set_engine(engine = "auto_arima_xgboost") %>%
    fit(Percent_change ~ date + as.numeric(date) + factor(month(date, label = TRUE), ordered = F),
        data = training(splits))

model_fit_ets <- exp_smoothing() %>%
    set_engine(engine = "ets") %>%
    fit(Percent_change ~ date, data = training(splits))

model_fit_prophet <- prophet_reg() %>%
    set_engine(engine = "prophet") %>%
    fit(Percent_change ~ date, data = training(splits))

model_fit_lm <- linear_reg() %>%
    set_engine("lm") %>%
    fit(Percent_change ~ as.numeric(date) + factor(month(date, label = TRUE), ordered = FALSE),
        data = training(splits))
```

```{r}
# Add the fitted models to a table
models_tbl <- modeltime_table(
    model_fit_arima_no_boost,
    model_fit_arima_boosted,
    model_fit_ets,
    model_fit_prophet,
    model_fit_lm
)
renderTable({
  models_tbl
})
```

We calibrate the model:

```{r}
calibration_tbl <- models_tbl %>%
    modeltime_calibrate(new_data = testing(splits))

renderTable({
  calibration_tbl
})
```

```{r}
renderPlot({
  calibration_tbl %>%
    modeltime_forecast(
        new_data    = testing(splits),
        actual_data = jalisco_work
    ) %>%
    plot_modeltime_forecast(
      .legend_max_width = 25
    )
})
```

```{r}
renderUI({
  calibration_tbl %>%
    modeltime_accuracy() %>%
    table_modeltime_accuracy()
})
```

```{r}
calibration_tbl

calibration_tbl %>%
    modeltime_forecast(
        new_data    = testing(splits),
        actual_data = jalisco_work
    ) %>%
    plot_modeltime_forecast(
      .legend_max_width = 25
    )

calibration_tbl %>%
    modeltime_accuracy() %>%
    table_modeltime_accuracy()
```

